---
title: "Data Accessibility Status"
author: "Joanna Colovas"
date: "November 21st 2023"
output: ioslides_presentation
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Project Goals
- Use mikropml (ML) R package to classify ASM papers as "containing new sequence data" or not based on most common "grams" (words) from the paper
- Report statistics on number of citations/paper as a function of data availability

## Current Status

- Still need to be done from Adena (20230731)
  + Complete testing of Random Forest(RF) model on classification of data availability (with tuning of hyperparameters)
  + Need to classify more papers to investigate false positives/false negatives from model, and appearances of duplicate papers


## Proposed Methodology
- Scrape ASM papers from Web of Science 
- Sort alphabetically by title to remove duplicates
- Create list of 'grams' for each paper
    - See if there's a way to keep common words/phrases together based on HTML (i.e *Vibrio cholerae*, Illumina MiSeq)
- Use mikropml to create a model of the most predictive words that aid in classifying papers **which model?**
   - Ensure feature table is generated, and comb the top X words by hand (that give model X% sensitivity/specificity)
   - Better to acknowledge that it's not going to be perfect but disclose error rate of Y% (10% error? 5%? Let model tell us?)

## Proposed Methodology
- Use generated lists from model to write script that will classify the rest of the papers from ASM
- Statistical analysis after paper classification, look at relationship between data availability and avg number of citations
   - Does this vary by ASM journal? 

  

