---
title: "2025079_dharma_plots"
output: html_document
date: "2025-07-09"
---

```{r setup, include=FALSE}
library(tidyverse)
library(MASS)
library(DHARMa)
library(sjPlot)
library(effects)
library(ggeffects)
```

```{r loading_data, include=FALSE}
#load local metadata
nsd_yes_metadata <- read_csv("~/Documents/Schloss/Colovas_Data_Accessibility/Data/final/nsd_yes_metadata.csv.gz")

#filter out NAs (important later)
nsd_yes_metadata <- nsd_yes_metadata %>%
  filter(., age.in.months != "NA" & da_factor != "NA" & container.title != "NA") %>% 
  mutate(da_factor = factor(da_factor), 
         container.title = factor(container.title))

#works faster locally
total_model <-glm.nb(is.referenced.by.count~ da_factor + log(age.in.months) + container.title + 
        + container.title*da_factor + log(age.in.months)*da_factor + container.title*log(age.in.months) + 
        log(age.in.months)*da_factor*container.title, data = nsd_yes_metadata, link = log)

```

## Using DHARMa package to evaluate the fit of our negative binomial model

-   Model format for all data from all journals **(N = 47353)**

    -   MASS::glm.nb(is.referenced.by.count\~ da_factor +
        log(age.in.months) + container.title +
        container.title\*da_factor + log(age.in.months)\*da_factor +
        container.title\*log(age.in.months) +
        log(age.in.months)\*da_factor\*container.title, data =
        nsd_yes_metadata, link = log)
    -   Data nsd_yes_metadata was filtered to remove all NAs from
        variables da_factor, age.in.months, and container.title to allow
        for some of the below visualizations.
    -   See below for number of papers from each journal.

    ```{r container.title.n, echo=FALSE}
    nsd_yes_metadata %>%  
      count(container.title, sort = T) %>% 
      knitr::kable()
    ```

## Using Predict to Graph the lines

```{r}
# predict.glm(total_model, type = "terms",)
plot_model(total_model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120,180]")) 


```

### Plot Residuals

-   Graph is very busy with all 47K observations

-   see 2nd plot for 5K observations for a smaller group to examine
    (index 25000-30000)

-   Residuals look like an amorphous blob as suggested by Abner\@CSCAR

```{r}
simulationOutput <- simulateResiduals(fittedModel = total_model, plot = F)
residuals(simulationOutput) %>% str()

#there is no alpha parameter in base R plot 
residuals(simulationOutput)  %>% 
  plot(main = "Residuals plotted by numerical index", pch = ".")

# residuals(simulationOutput)[25000:30000] %>% 
#   plot(main = "Residuals for observation indices 25000 - 3000")
```

## Plots for Residuals - QQ and DHARMa

### QQ Plot

-   KS Test = Two sample Kolmogorov-Smirnov (KS) Test

    -   This function tests the overall uniformity of the simulated
        residuals in a DHARMa object - Deviation is significant between
        the expected residuals and the actual observed residuals.

    -   "If the P value is small, conclude that the two groups were
        sampled from populations with different distributions. " -Prism
        help page

-   Dispersion Test

    -   This function performs simulation-based tests for
        over/underdispersion

    -   Over / underdispersion means that the observed data is more /
        less dispersed than expected under the fitted model.

    -   Deviation is significant between the observed data and fitted
        model.

-   Outlier Test

    -   This function tests if the number of observations outside the
        simulation envelope are larger or smaller than expected

    -   Methods generate a null expectation, and then test for an excess
        or lack of outliers. Per default, testOutliers() looks for both,
        so if you get a significant p-value, you have to check if you
        have to many or too few outliers.

    -   See Outlier test for distribution of outliers. - Many at 1.0
        residual.

```{r}
plot(simulationOutput, sub = "All data glm(fixed)")
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
DHARMa::testOutliers(simulationOutput, type = "bootstrap")
```

## Looking at each model by journal

```{r}
#smaller model with 2 terms
two_term_glmnb <-function(model_data, model_name) {

  total_model <-MASS::glm.nb(is.referenced.by.count~ da_factor + log(age.in.months) + 
       + log(age.in.months)*da_factor + log(age.in.months)*da_factor, data = model_data, link = log)

  return(total_model)
 
}

journals <- 
  nsd_yes_metadata %>%  
  count(journal_abrev) %>%  
  filter(journal_abrev != "jmbe") 

#j <- 5

for(j in 1:nrow(journals)) { 
  journal_data <- 
  nsd_yes_metadata %>% 
    filter(journal_abrev == journals[[j,1]]) %>%  
    mutate(da_factor = factor(da))


actual_citations <- 
 journal_data %>% 
    filter(age.in.months %in% c(12, 60, 84, 120, 180)) %>% 
    group_by(age.in.months, da_factor) %>%  
   count(is.referenced.by.count, .drop = FALSE) %>% 
   summarize(mean_citations = mean(is.referenced.by.count)) %>% 
  ggplot(aes(x = da_factor, y = mean_citations, color = factor(age.in.months))) + 
  geom_point() + 
  ggtitle(paste0("Observed mean citations\nPlotted for journal ", journals[[j,1]]))
print(actual_citations)

  model <- two_term_glmnb(journal_data, journals[[j,1]])
  
#  p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", 
  # "age.in.months[12,60,84,120,180]"))
  


  
  if(j == 3) { #genomea = 5
  plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[84,120,144]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]]))
  
  p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[84,120,144]"))
  }
  else if (j == 9) { ##Mra - 7 
    plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]]))
    p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84]"))
    
  }
  else if (j == 10 | j == 11 ) { #msphere, msystems 9 years
  plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,108]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]]))
  
  p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,108]"))
  }
  else if (j == 12 ) { 
      plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]]))
      p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120]"))
  }
  else { 
     plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120,180]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]]))
     p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120,180]"))
  }
  
    ratio_plot <- tibble(da = p$x, predicted = p$predicted, age.in.months = as.numeric(as.character(p$group))) %>%  
    pivot_wider(names_from = da, values_from = predicted) %>%  
    mutate(ratio = `2`/`1`) %>%  
    ggplot(., aes(x = age.in.months, y = ratio)) + 
     ratio_plot <- 
      tibble(da = p$x, predicted = p$predicted, age.in.months = as.numeric(as.character(p$group)), low = p$conf.low, high = p$conf.high) %>%  
    pivot_wider(names_from = da, values_from = c(predicted, low, high)) %>%  
    mutate(high_high_ratio = `high_2`/`high_1`,
           high_mid_ratio = `high_2`/`predicted_1`,
           high_low_ratio = `high_2`/`low_1`,
           high_high_ratio = `high_2`/`high_1`,
           mid_high_ratio = `predicted_2`/`high_1`,
           mid_mid_ratio = `predicted_2`/`predicted_1`,
           mid_low_ratio = `predicted_2`/`low_1`, 
           low_high_ratio = `low_2`/`high_1`, 
           low_mid_ratio = `low_2`/`predicted_1`,
           low_low_ratio = `low_2`/`low_1`,
           ) %>%  
        dplyr::select(., age.in.months, ends_with("ratio")) %>% 
        pivot_longer(., cols = ends_with("ratio"), names_to = "ratio_name", 
                     values_to = "ratio_value") %>% 
    ggplot(., aes(x = age.in.months, y = ratio_value)) + 
    geom_boxplot(aes(group = age.in.months)) +
    geom_point() + 
     labs(title = paste0("Ratio of predicted citations for da = Yes to da = No \nPlotted for journal ", journals[[j,1]]), 
          y = "Ratio number of citations da=Yes/da=No")
    
 print(plot_model)
 print(ratio_plot)
 
  simulation <- simulateResiduals(fittedModel = model, plot = F)
  
  # residuals(simulation)  %>% 
  # plot(main = paste0("Residuals plotted for journal ", journals[[j,1]]), pch = ".")
  residuals <- 
   residuals(simulation) %>%  tibble(residual = .) %>%  
     ggplot(aes(y = residual, x = seq_along(residual) )) + 
     geom_point(size = 1) + 
     geom_smooth() +
     ggtitle(paste0("Residuals with geom_smooth line' \nPlotted for journal ", journals[[j,1]]))
  print(residuals)
  
   residuals_hist <-
   residuals(simulation) %>%  tibble(residual = .) %>%  
     ggplot(aes(x = residual)) + 
     geom_histogram(linewidth = 0.25, color = "white", binwidth = 1/100) + 
     ggtitle(paste0("Residuals Plotted for journal ", journals[[j,1]]))
  print(residuals_hist)
  
  plot(simulation, sub = paste0("Simulation plotted for journal ", journals[[j,1]]))
  
  
  }
```

## JB 10 years

```{r}
 j <-  5
  journal_data <- 
  nsd_yes_metadata %>% 
    filter(journal_abrev == journals[[j,1]]) %>%  
    mutate(da_factor = factor(da)) %>%  
    filter(age.in.months <= 120)
  


actual_citations <- 
 journal_data %>% 
    filter(age.in.months %in% c(12, 60, 84, 120)) %>% 
    group_by(age.in.months, da_factor) %>%  
   count(is.referenced.by.count, .drop = FALSE) %>% 
   summarize(mean_citations = mean(is.referenced.by.count)) %>% 
  ggplot(aes(x = da_factor, y = mean_citations, color = factor(age.in.months))) + 
  geom_point() + 
  ggtitle(paste0("Observed mean citations\nPlotted for journal ", journals[[j,1]]))
print(actual_citations)

  model <- two_term_glmnb(journal_data, journals[[j,1]])
  # assign(journal_abrev[[j,1]], model)
  
  
     plot_model <- plot_model(model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120]")) +     ggtitle(paste0("Predicted counts of 'is.referenced.by.count' \nPlotted for journal ", journals[[j,1]], " for the last 10 years"))
  
  p <-  get_model_data(model = model, type = "pred", terms = c("da_factor", "age.in.months[12,60,84,120,180]"))
  
  ratio_plot <- tibble(da = p$x, predicted = p$predicted, age.in.months = as.numeric(as.character(p$group))) %>%  
    pivot_wider(names_from = da, values_from = predicted) %>%  
    mutate(ratio = `2`/`1`) %>%  
    ggplot(., aes(x = age.in.months, y = ratio)) + 
 ratio_plot <- 
      tibble(da = p$x, predicted = p$predicted, age.in.months = as.numeric(as.character(p$group)), low = p$conf.low, high = p$conf.high) %>%  
    pivot_wider(names_from = da, values_from = c(predicted, low, high)) %>%  
    mutate(high_high_ratio = `high_2`/`high_1`,
           high_mid_ratio = `high_2`/`predicted_1`,
           high_low_ratio = `high_2`/`low_1`,
           high_high_ratio = `high_2`/`high_1`,
           mid_high_ratio = `predicted_2`/`high_1`,
           mid_mid_ratio = `predicted_2`/`predicted_1`,
           mid_low_ratio = `predicted_2`/`low_1`, 
           low_high_ratio = `low_2`/`high_1`, 
           low_mid_ratio = `low_2`/`predicted_1`,
           low_low_ratio = `low_2`/`low_1`,
           ) %>%  
        dplyr::select(., age.in.months, ends_with("ratio")) %>% 
        pivot_longer(., cols = ends_with("ratio"), names_to = "ratio_name", 
                     values_to = "ratio_value") %>% 
    ggplot(., aes(x = age.in.months, y = ratio_value)) + 
    geom_boxplot(aes(group = age.in.months)) +
    geom_point() + 
     labs(title = paste0("Ratio of predicted citations for da = Yes to da = No \nPlotted for journal ", journals[[j,1]]), 
          y = "Ratio number of citations da=Yes/da=No")

 print(plot_model)
 print(ratio_plot)
 
  simulation <- simulateResiduals(fittedModel = model, plot = F)
  
  # residuals(simulation)  %>% 
  # plot(main = paste0("Residuals plotted for journal ", journals[[j,1]]), pch = ".")
  
   residuals(simulation) %>%  tibble(residual = .) %>%  
     ggplot(aes(y = residual, x = seq_along(residual) )) + 
     geom_point(size = 1) + 
     geom_smooth() +
     ggtitle(paste0("Residuals with geom_smooth line' \nPlotted for journal ", journals[[j,1]], " for the last 10 years"))
  
  plot(simulation, sub = paste0("Simulation plotted for journal ", journals[[j,1]]))
  
  
     residuals_hist <-
   residuals(simulation) %>%  tibble(residual = .) %>%  
     ggplot(aes(x = residual)) + 
     geom_histogram(linewidth = 0.25, color = "white", binwidth = 1/100) + 
     ggtitle(paste0("Residuals Plotted for journal ", journals[[j,1]]))
  print(residuals_hist)
  
  plot(simulation, sub = paste0("Simulation plotted for journal ", journals[[j,1]]))
  

```
