---
title: "Negative Binomial Regression Fit Report"
output: pdf_document
date : 20250610 
classoption: landscape
---

## Project Summary

-   We are using data from the American Society of Microbiology's (ASM)
    12 published journals to investigate the relationship between the
    number of citations (variable 'is.referenced.by.count') a published
    scientific article receives and if the authors have included access
    to their raw sequencing data (variable 'da', data availability) in
    the manuscript.
-   We are trying to understand if publishing raw data helps to improve
    citation metrics. We have data from 2000-2024, and will also adjust
    for time published (variable 'age.in.months'), as older papers have
    had the opportunity to accumulate more citations over time.

```{r include=FALSE}
library(tidyverse)

# library(MASS)
# library(jtools)

rsquared <- read_csv("~/Documents/Schloss/Colovas_Data_Accessibility/Data/final/negative_binomial_models.csv") %>%  
  rename(all_data_rsq = all_journal_data_rsq)

all_journals <- read_csv("~/Documents/Schloss/Colovas_Data_Accessibility/Data/negative_binomial/all_data_glmnb_models.csv")

each_journal <-  read_csv("~/Documents/Schloss/Colovas_Data_Accessibility/Data/negative_binomial/negative_binomial_byjournal.csv.gz") %>%  
  rename(all_journal_value = journal_value)

```

```{r}

knitr::kable(rsquared, digits = 4)
```

## How well do the models fit (by Cragg-Uhler pseduo R-squared metric)

-   See above table "rsquared"

-   Model format for all data from all journals
    MASS::glm.nb(is.referenced.by.count\~ da_factor +
    log(age.in.months) + container.title + container.title\*da_factor +
    log(age.in.months)\*da_factor +
    container.title\*log(age.in.months) +
    log(age.in.months)\*da_factor\*container.title, data =
    nsd_yes_metadata, link = log)

-   Use model format for data from each journal
    MASS::glm.nb(is.referenced.by.count\~ da_factor +
    log(age.in.months) + log(age.in.months)\*da_factor, data = \<each
    journal\>, link = log)

-   **Overall model fit with all data from all journals:**

    -   R\^2 value = 0.678
    -   Removal of top 1% of data: R\^2 value = 0.682
    -   Truncate data to last 5 years: R\^2 value = 0.660
    -   Truncate data to last 10 years: R\^2 value = 0.680
    -   **Summary :** Model fit by R\^2 metric does not change by
        removing the top 1% of data or truncating to data from the last
        5 or 10 years.

-   **Overall model fit for data from EACH journal individually:**

    -   4/12 journals have **overall model fit** with R\^2 \> 0.5

    -   4/12 journals have fit with R\^2 \> 0.5 with **top 1% of data
        removed**

    -   10/11 journals have model fits \>0.5 when **truncated to the
        last 5 years,** so they are better than their fit overall (one
        journal has no data from this period)

    -   8/12 journals have model fits \>0.5 when **truncated to the last
        10 years,** so they are better than their fit overall

    -   **Summary:** Data fits negative binomial model better with only
        more recent data considered.

```{r}
knitr::kable(all_journals, digits = 4) 
```

## All journal model is resistant to changes from removing top 1% of data, but less resistant to changes from truncating at 5 and 10 years.

-   When working across the columns in the second table, we have
    coefficients on the left, followed by their values under the
    following conditions

    -   full_model_value = all data included in the model
    -   no_1percent_value = top 1% of data removed
    -   five_years_value = data truncated at 5 years in age of paper
    -   ten_years_value = data truncated at 10 years in age of paper
    -   **Note:** Journal of Microbiology and Biology Education(jmbe)
        has N=7 papers with new sequence data and has been excluded for
        these analyses, but is a part of the model, and appears as NAs
        in the table above.

-   Would graphing be helpful here to look at the ways the values change
    over these transformations? What kind of graphing?

```{r}
knitr::kable(each_journal, digits = 4)
```

## Each journal model are semi-resistant to changes from removing top 1% of data, and even less resistant to changes from truncating at 5 and 10 years.

-   See above for mutations on these columns, but these models look less
    resistant to the transformation of removing the top 1% of data, and
    even less resistant to changes in coefficients from truncating at 5
    and 10 years of data.

-   Would graphing be helpful here to look at the ways the values change
    over these transformations? What kind of graphing?
